A Gaussian process model of human e\-lect\-ro\-cort\-i\-co\-graph\-ic data

### 2020-03-11 paper/main.tex
updated supplemental figure references in main text, added references to S7 and S8

\newcommand{\perexpcorrmaps}{S1}
\newcommand{\perexptaskrecon}{S2}
\newcommand{\perexptaskreconseparated}{S3}
\newcommand{\freqpower}{S4}
\newcommand{\networkpower}{S5}
\newcommand{\suppstats}{S6}

@@

\section*{Results}
We used a cross-validation approach to test the accuracy with
which the SuperEEG algorithm reconstructs activity throughout the brain. For
each patient in turn, we estimated full-brain correlation matrices
(Eqn.~\ref{eqn:corrmat}) using data from all of the \textit{other} patients.
This step ensured that the data we were reconstructing could not also be used to
estimate the between-location correlations that drove the reconstructions via
Equation~\ref{eqn:reconstruction} (otherwise the analysis would be circular).
For that held-out patient, we held out each electrode in turn.  We used
Equation~\ref{eqn:reconstruction} to reconstruct activity at the held-out
electrode location, using the correlation matrix learned from all other
patients' data as $\hat{K}$, and using activity recorded from the other
electrodes from the held-out patient as $Y_{s, k, \alpha_s}$.  We then asked:
how closely did each of the SuperEEG-estimated recordings at those electrodes
match the observed recordings from those electrodes (i.e., how closely did the
estimated $\hat{Y}_{s, k, \beta_s}$ match the observed $Y_{s, k, \beta_s}$)?

@@

Next, we compared the quality of these across-participant reconstructions (i.e.,
computed using a correlation model learned from other patients' data) to
reconstructions generated using a correlation model trained using the
in-patient's data.  In other words, for this within-patient benchmark analysis
we estimated $\hat{C}_{s}$ (Eqn.~\ref{eqn:subj_corrmat}) for each patient in
turn, using recordings from all of that patient's electrodes except at the
location we were reconstructing.  These within-patient reconstructions serve as
an estimate of how well data from all of the other electrodes from that single
patient may be used to estimate held-out data from the same patient.  This
allows us to ask how much information about the activity at a given electrode
might be inferred through (a) volume conductance or other sources of ``leakage''
from activity patterns measured from the patient's other electrodes and (b)
across-electrode correlations learned from that single patient.  As shown in
Figure~\ref{fig:corrmap}A (gray histogram), the distribution of within-patient
correlations was centered well above zero (mean: $r = 0.32$; $t$-test comparing
mean of distribution of $z$-transformed average patient correlation coefficients
to 0: $t(66) = 15.16, p < 10^{-10}$). However, the across-patient correlations
were substantially higher ($t$-test comparing average $z$-transformed within
versus across patient electrode correlations: $t(66) = 9.17, p < 10^{-10}$).
This is an especially conservative test, given that the across-patient SuperEEG
reconstructions exclude (from the correlation matrix estimates) all data from
the patient whose data is being reconstructed.  We repeated each of these
analyses on a second independent dataset and found similar results
(Fig.~\ref{fig:corrmap}B, D; within versus across reconstruction accuracy:
$t(77) = 11.25, p < 10^{-10}$). We also replicated this result separately for
each of the two experiments from Dataset 2 (Fig.~\perexpcorrmaps).  This overall
finding, that reconstructions of held-out data using correlation models learned
from \textit{other} patient's data yield higher reconstruction accuracy than
correlation models learned from the patient whose data is being reconstructed,
has two important implications.  First, it implies that distant electrodes
provide additional predictive power to the data reconstructions beyond the
information contained solely in nearby electrodes.  This follows from the fact
that each patient's grid, strip, and depth electrodes are implanted in a unique
set of locations, so for any given electrode the closest electrodes in the full
dataset tend to come from the same patient.  Second, it implies that the
spatial correlations learned using the SuperEEG algorithm are, to some extent,
similar across people.

%correlations by brain area, distribution of correlations
\begin{figure}
  \centering \includegraphics[width=\textwidth]{figs/corrmap}
  \caption{\textbf{Reconstruction accuracy across all electrodes in two ECoG
  datasets.}  \textbf{A. Distributions of correlations between observed versus
  reconstructed activity by electrode, for Dataset 1.}  The across-patient
  distribution (black) reflects reconstruction accuracy (correlation) using a
  correlation model learned from all but one patient's data, and then applied to
  that held-out patient's data.  The within-patient distribution (gray) reflects
  performance using a correlation model learned from the same patient who
  contributed the to-be-reconstructed electrode. \textbf{B. Distributions of
  correlations for Dataset 2.}  This panel is in the same format as Panel A, but
  reflects results obtained from Dataset 2.  The histograms aggregate data
  across both Dataset 2 experiments; for results broken down by experiment see
  Figure~\perexptaskreconseparated. \textbf{C.--D.  Reconstruction accuracy by
  location.} The colors denote the average across-session correlations, using
  the across-patient correlation model, between the observed and reconstructed
  activity at the given electrode location projected to the cortical
  surface~\citep{CombEtal19}.  Panel C displays the map for Dataset 1 and Panel
  D displays the map for Dataset 2.} \label{fig:corrmap}
\end{figure}

@@

To test the task-specificity of the SuperEEG-derived correlation models, we
restricted the dataset to the 24 patients that participated in both experiments
and repeated the above within- and across-patient cross validation procedures
separately for Experiment 1 and Experiment 2 data from Dataset 2.  We then
compared the reconstruction accuracies for held-out electrodes, for models
trained within versus across the two experiments, or combining across both
experiments (Fig.~\perexptaskrecon).  In every case we found that across-patient
models trained using data from all other patients out-performed within-patient
models trained on data only from the subject contributing the given electrode
($t$s$(23) > 6.50, p$s$ < 10^{-5}$). All reconstruction accuracies also reliably
exceeded chance performance ($t$s$(23) > 8.00, p$s$ < 10^{-8}$).  Average
reconstruction accuracy was highest for the across-patient models limited to
data from the same experiment (mean accuracy: $r = 0.68$); next-highest for the
models that combined data across both experiments (mean accuracy: $r = 0.61$); and
lowest for models trained across tasks (mean accuracy: $r = 0.47$).  This pattern of
results also held for each of the Dataset 2 experiments individually
(Fig.~\perexptaskreconseparated).  Taken together, these results indicate that
there are reliable commonalities in the spatial correlations of full-brain
activity across tasks, but that there are also reliable differences in these
spatial correlations across tasks. Whereas reconstruction accuracy benefits from
incorporating data from other patients, reconstruction accuracy is highest when
constrained to within-task data, or data that includes a variety of tasks (e.g.,
Dataset 1, or combining across the two Dataset 2 experiments).

@@

To gain additional insight into which regions were most informative about
full-brain activity patterns at different frequency bands, we next computed (for
each frequency band) the intersection of the top 10\% highest information scores
across the maps for Datasets 1 and 2 (analogous to our approach in
Fig.~\ref{fig:informap}C). This yielded a single map of the (reliably) most
informative locations, for each frequency band we examined. We then carried out
\textit{post hoc} analyses on each of these maps to characterize the underlying
structural and functional properties of each set of regions we identified as
being particularly informative about one or more types of neural pattern.

>>>>

\newcommand{\combinedexpviolins}{S1}
\newcommand{\splitexpviolins}{S2}
\newcommand{\splitexpcorrmaps}{S3}
\newcommand{\freqpower}{S4}
\newcommand{\networkpower}{S5}
\newcommand{\suppstats}{S6}
\newcommand{\temporalstability}{S7}
\newcommand{\patientstability}{S8}

@@

\section*{Results}
We used a cross-validation approach to test the accuracy with which
the SuperEEG algorithm reconstructs activity throughout the brain. For
each patient in turn, we estimated full-brain correlation matrices
(Eqn.~\ref{eqn:corrmat}) using data from all of the \textit{other}
patients.  This step ensured that the data we were reconstructing
could not also be used to estimate the between-location correlations
that drove the reconstructions via Equation~\ref{eqn:reconstruction}
(otherwise the analysis would be circular).  For that held-out
patient, we held out each electrode in turn.  We used
Equation~\ref{eqn:reconstruction} to reconstruct activity at the
held-out electrode location, using the correlation matrix learned from
all other patients' data as $\hat{K}$, and using activity recorded
from the other electrodes from the held-out patient as
$Y_{s, k, \alpha_s}$.  (For analyses examining the stability of our
estimates of $\hat{K}$ across time and patients, see
Figs.~\temporalstability~and \patientstability, respectively).  We then asked: how closely did each of the
SuperEEG-estimated recordings at those electrodes match the observed
recordings from those electrodes (i.e., how closely did the estimated
$\hat{Y}_{s, k, \beta_s}$ match the observed $Y_{s, k, \beta_s}$)?

@@

Next, we compared the quality of these across-participant reconstructions (i.e.,
computed using a correlation model learned from other patients' data) to
reconstructions generated using a correlation model trained using the
in-patient's data.  In other words, for this within-patient benchmark analysis
we estimated $\hat{C}_{s}$ (Eqn.~\ref{eqn:subj_corrmat}) for each patient in
turn, using recordings from all of that patient's electrodes except at the
location we were reconstructing.  These within-patient reconstructions serve as
an estimate of how well data from all of the other electrodes from that single
patient may be used to estimate held-out data from the same patient.  This
allows us to ask how much information about the activity at a given electrode
might be inferred through (a) volume conductance or other sources of ``leakage''
from activity patterns measured from the patient's other electrodes and (b)
across-electrode correlations learned from that single patient.  As shown in
Figure~\ref{fig:corrmap}A (gray histogram), the distribution of within-patient
correlations was centered well above zero (mean: $r = 0.32$; $t$-test comparing
mean of distribution of $z$-transformed average patient correlation coefficients
to 0: $t(66) = 15.16, p < 10^{-10}$). However, the across-patient correlations
were substantially higher ($t$-test comparing average $z$-transformed within
versus across patient electrode correlations: $t(66) = 9.17, p < 10^{-10}$).
This is an especially conservative test, given that the across-patient SuperEEG
reconstructions exclude (from the correlation matrix estimates) all data from
the patient whose data is being reconstructed.  We repeated each of these
analyses on a second independent dataset and found similar results
(Fig.~\ref{fig:corrmap}B, D; within versus across reconstruction accuracy:
$t(77) = 11.25, p < 10^{-10}$). We also replicated this result separately for
each of the two experiments from Dataset 2 (Fig.~\splitexpcorrmaps).  This overall
finding, that reconstructions of held-out data using correlation models learned
from \textit{other} patient's data yield higher reconstruction accuracy than
correlation models learned from the patient whose data is being reconstructed,
has two important implications.  First, it implies that distant electrodes
provide additional predictive power to the data reconstructions beyond the
information contained solely in nearby electrodes.  This follows from the fact
that each patient's grid, strip, and depth electrodes are implanted in a unique
set of locations, so for any given electrode the closest electrodes in the full
dataset tend to come from the same patient.  Second, it implies that the
spatial correlations learned using the SuperEEG algorithm are, to some extent,
similar across people.

%correlations by brain area, distribution of correlations
\begin{figure}
  \centering \includegraphics[width=\textwidth]{figs/corrmap}
  \caption{\textbf{Reconstruction accuracy across all electrodes in two ECoG
  datasets.}  \textbf{A. Distributions of correlations between observed versus
  reconstructed activity by electrode, for Dataset 1.}  The across-patient
  distribution (black) reflects reconstruction accuracy (correlation) using a
  correlation model learned from all but one patient's data, and then applied to
  that held-out patient's data.  The within-patient distribution (gray) reflects
  performance using a correlation model learned from the same patient who
  contributed the to-be-reconstructed electrode. \textbf{B. Distributions of
  correlations for Dataset 2.}  This panel is in the same format as Panel A, but
  reflects results obtained from Dataset 2.  The histograms aggregate data
  across both Dataset 2 experiments; for results broken down by experiment see
  Figures~\splitexpviolins~and~\splitexpcorrmaps. \textbf{C.--D.  Reconstruction accuracy by
  location.} The colors denote the average across-session correlations, using
  the across-patient correlation model, between the observed and reconstructed
  activity at the given electrode location projected to the cortical
  surface~\citep{CombEtal19}.  Panel C displays the map for Dataset 1 and Panel
  D displays the map for Dataset 2.} \label{fig:corrmap}
\end{figure}

@@

To test the task-specificity of the SuperEEG-derived correlation models, we
restricted the dataset to the 24 patients that participated in both experiments
and repeated the above within- and across-patient cross validation procedures
separately for Experiment 1 and Experiment 2 data from Dataset 2.  We then
compared the reconstruction accuracies for held-out electrodes, for models
trained within versus across the two experiments, or combining across both
experiments (Fig.~\combinedexpviolins).  In every case we found that across-patient
models trained using data from all other patients out-performed within-patient
models trained on data only from the subject contributing the given electrode
($t$s$(23) > 6.50, p$s$ < 10^{-5}$). All reconstruction accuracies also reliably
exceeded chance performance ($t$s$(23) > 8.00, p$s$ < 10^{-8}$).  Average
reconstruction accuracy was highest for the across-patient models limited to
data from the same experiment (mean accuracy: $r = 0.68$); next-highest for the
models that combined data across both experiments (mean accuracy: $r = 0.61$); and
lowest for models trained across tasks (mean accuracy: $r = 0.47$).  This pattern of
results also held for each of the Dataset 2 experiments individually
(Fig.~\splitexpviolins).  Taken together, these results indicate that
there are reliable commonalities in the spatial correlations of full-brain
activity across tasks, but that there are also reliable differences in these
spatial correlations across tasks. Whereas reconstruction accuracy benefits from
incorporating data from other patients, reconstruction accuracy is highest when
constrained to within-task data, or data that includes a variety of tasks (e.g.,
Dataset 1, or combining across the two Dataset 2 experiments).

@@

To gain additional insight into which regions were most informative about
full-brain activity patterns at different frequency bands, we next computed (for
each frequency band) the intersection of the top 10\% highest information scores
across the maps for Datasets 1 and 2 (analogous to our approach in
Fig.~\ref{fig:informap}C). This yielded a single map of the (reliably) most
informative locations, for each frequency band we examined. We then carried out
\textit{post hoc} analyses on each of these maps to characterize the underlying
structural and functional properties of each set of regions we identified as
being particularly informative about one or more types of neural
pattern (Figs.~\ref{fig:networks} and~\networkpower).
